An Investigation of Confusing Code Patterns in JavaScript

### 2023-01-11 conclusion.tex
Added file response_reviewer_1_2nd_round.md that addresses all the comments made by reviewer 1. I have also changed the order of the studies and made the appropriate adaptations for the MSR study to appear first, before the others and as a motivation for them.

This paper reports the results of a mixed-method research effort that allowed us to better understand the impact of atoms of confusion in JavaScript code. First, we conducted two independent experiments, which involve 210 JavaScript developers in total, asking them to predict the output of some code snippets (with and without atoms candidates). In at least one of the experiments, five atom candidates (Type Conversion, Change of Literal Encoding, Omitted Curly Braces, Pre-Increment, and Post- Increment) significantly impact program comprehension activities and can be considered atoms of confusion.
%% : Comma Operator, Automatic Semicolon Insertion, Post Increment, Assignment as Value, and Implicit Predicate. 
%% We also interviewed 15 professional developers. In general, they considered code without atom candidates easier to understand and highlighted other JavaScript constructs and idioms that might also introduce misunderstanding. 
%Last, but not least,
%Finally, we conducted a mining software repositories study. We found four atom candidates frequently used in JavaScript code, i.e., Implicit Predicate, Post Increment, Ternary Operator, and Omitted Curly Braces.  %The first two can be considered atoms of confusion. 
Our efforts have two implications for practice. First, we present evidence that not all atoms of confusion validated to \clang, \cpplang, or Java led to a statistically significant impact on program understanding for JavaScript. Besides that, our findings might be used to alert developers to avoid writing JavaScript code with certain atoms of confusion (e.g., Comma Operator, Automatic Semicolon Insertion, Post Increment/Decrement, Assignment as Value, and Implicit Predicate). Finally, our results might help tool developers to create program transformation tools for removing atoms that frequently appear in software. 
%They might firstly focus, for instance, on the atoms we found to be more confusing and at the same time more common in practice (e.g., Omitted Curly Braces and Post Increment).
%Promising candidates may be the atoms we found to be most confusing and, at the same time, common in practice (e.g., Implicit Predicate and Post Increment).
%As mentioned, the results of the interviews
Our research also pointed out additional JavaScript constructs and idioms 
that might introduce misunderstanding. These include nested callbacks, % and prototype-based inheritance (which might not have a counterpart version), but also idioms such
%as 
\emph{property access}, and \emph{arrow functions}. As future work, we intend to
reproduce our survey to validate whether or not these additional
idioms 
%, which the interviewees pointed out, 
truly affect
the understanding of JavaScript code.

>>>>

This paper reports the results of a mixed-method research effort that allowed us to better understand the impact of atoms of confusion in JavaScript code. This involved a repository mining study, two experiments, and an interview study. The two independently-conducted experiments, which involved 210 JavaScript developers in total, asked them to predict the output of code snippets with and without atoms candidates. In at least one of the experiments, five atom candidates (Type Conversion, Change of Literal Encoding, Omitted Curly Braces, Pre-Increment, and Post- Increment) significantly impact program comprehension activities and can be considered atoms of confusion.
%% : Comma Operator, Automatic Semicolon Insertion, Post Increment, Assignment as Value, and Implicit Predicate. 
%% We also interviewed 15 professional developers. In general, they considered code without atom candidates easier to understand and highlighted other JavaScript constructs and idioms that might also introduce misunderstanding. 
%Last, but not least,
%Finally, we conducted a mining software repositories study. We found four atom candidates frequently used in JavaScript code, i.e., Implicit Predicate, Post Increment, Ternary Operator, and Omitted Curly Braces.  %The first two can be considered atoms of confusion. 
Our efforts have two implications for practice. First, we present evidence that not all atoms of confusion validated to \clang, \cpplang, or Java led to a statistically significant impact on program understanding for JavaScript. Besides that, our findings might be used to alert developers to avoid writing JavaScript code with certain atoms of confusion (e.g., Comma Operator, Automatic Semicolon Insertion, Post Increment/Decrement, Assignment as Value, and Implicit Predicate). Finally, our results might help tool developers to create program transformation tools for removing atoms that frequently appear in software. 
%They might firstly focus, for instance, on the atoms we found to be more confusing and at the same time more common in practice (e.g., Omitted Curly Braces and Post Increment).
%Promising candidates may be the atoms we found to be most confusing and, at the same time, common in practice (e.g., Implicit Predicate and Post Increment).
%As mentioned, the results of the interviews
Our research also pointed out additional JavaScript constructs and idioms 
that might introduce misunderstanding. These include nested callbacks, % and prototype-based inheritance (which might not have a counterpart version), but also idioms such
%as 
\emph{property access}, and \emph{arrow functions}. As future work, we intend to
reproduce our survey to validate whether or not these additional
idioms 
%, which the interviewees pointed out, 
truly affect
the understanding of JavaScript code.

### 2023-01-11 introduction.tex
Added file response_reviewer_1_2nd_round.md that addresses all the comments made by reviewer 1. I have also changed the order of the studies and made the appropriate adaptations for the MSR study to appear first, before the others and as a motivation for them.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% INICIO INTRODUCAO - MARCIO
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Source code misunderstandings can occur due to many reasons; for instance, the wrong choice of names for arguments, variables, and methods might hinder developers from understanding a piece of code~\cite{avidan:icpc2017}. Besides that, programming languages can also have specific idioms and constructs that may lead to misunderstandings. That is the case of atoms of confusion: small pieces of code that prevent developers from program comprehension, even though functionally equivalent alternatives exist and are easier to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17, DBLP:conf/msr/GopsteinZFC18}.

Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ} has explored the impact of atoms of confusion on program understanding, and confirmed they negatively impact source code readability. For instance, consider the pair of JavaScript code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}). The code at the left-hand side contains two atoms of confusion discussed in previous work focusing on the \clang language~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: Logic as Control Flow and Post-increment. At the right-hand side, we present a clean version of the code, that is, without atoms of confusion. According to the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms Logic as Control Flow and Post-increment improved the accuracy of program understanding tasks in 41\% and 34\%, respectively.

\begin{figure*}[thb]
\noindent\begin{minipage}{.45\textwidth}  
\begin{lstlisting}[language=JavaScript]
res = properties.reduce((s, b) => {
  b && s++
  return s
}
\end{lstlisting}
\end{minipage}\hfill
\begin{minipage}{.45\textwidth}
\begin{lstlisting}[language=JavaScript]
res = properties.reduce((s, b) => {
  if(b) {
    s = s + 1
  }
  return s
}
\end{lstlisting}
\end{minipage}
\captionof{lstlisting}{Code snippet from \textsc{NervJS/taro} project. Line 2 on the left-hand side of the figure contains two
  atoms of confusion discussed in ~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: Logic
  as Control Flow and Post-increment. On the right-hand side, we present an alternative implementation without atoms of confusion. We could simplify the code on the left-hand side even more, by using \texttt{filter} and \texttt{length} instead of the \texttt{reduce} recursive pattern (e.g., \texttt{res = properties.filter(p => p.length})).}\label{fig:lst01}
\end{figure*}

Nonetheless, these previous studies focus exclusively on statically typed languages (e.g., \clang, \cpplang, and Java). This way, there is a knowledge gap with respect to the impact of such atoms in dynamic languages such as JavaScript. In this sense, because JavaScript's programming culture differs from that of languages such as \clang and \cpplang, particularly due to the JavaScript's dynamic capabilities and weaker type system, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa. Also, besides being the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}, JavaScript contains features that combined make the source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}.
%, such as weakly typed language~\cite{DynamicBehaviorJS:PLDI}, object-oriented, functional,
% and imperative constructs~\cite{JS-CodeSmells:SCAM}, and features such as run-time
% evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations.

In this way, to better understand the impact of atoms of confusion in JavaScript, in this paper we present a mixed-method research effort based on a survey, on interviews, and on a mining software repositories study. First, we surveyed 140 JavaScript developers. In the survey, their task was to predict the output of small code snippets, with half of them containing atoms of confusion. The number of correct answers for the \emph{clean version} of the code snippets for 7 out of \na atom candidates is at least 15\% higher than the corresponding \emph{confusing version} of the code snippet (i.e., with an atom candidate). For 5 of these atoms, the difference is statistically significant (accounting for p-value correction). Effect size is large for 4 atoms and medium for 1 of them. 
The presence of an atom candidate might also affect the time necessary for submitting a correct answer. For instance, the presence of the Comma Operator atom increases the average time to predict the output by 76.23\%, though the outputs of snippets containing the Pre-Increment atom were correctly predicted 38.19\% faster.
To learn more about the misunderstandings in the snippets used in the survey, we also interviewed 15 experienced professional developers. Participants prefered the code snippets without the atom candidates in 70\% of the cases, therefore corroborating the survey findings. The participants also discussed other idioms and constructs that might lead developers to misunderstand JavaScript code.

The frequency in which the atoms appear \emph{in the wild} might reveal cases where developers should be more or less careful when writing JavaScript code. To understand the incidence of atoms, we mined popular JavaScript projects from GitHub and found that atoms frequently appear in practice. 
Our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems. Four atom candidates appear in 90\% of the analyzed projects and two of them, Implicit Predicate and Ternary Operator, occur more than once for every 100 lines of code in the dataset. 
%Indeed, four atom candidates are frequently used in JavaScript code---the Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.

Altogether, the main contributions of this paper are:

\begin{itemize}

\item A mixed-method research effort---based on a survey, interviews, and mining software
  repositories---about the impact of atoms of confusion on JavaScript (a
    dynamic language).

  \item A comparison of the impact and prevalence of atoms of confusion in
    JavaScript with what has been previously reported for other languages
    (\clang, \cpplang, and Java).

    \item A library and infrastructure to mine atoms of confusion using 
    code queries (implemented in the .QL language). This library and 
    infrastructure are publicly available and might help researchers and 
    practitioners to automate program analysis at a large scale.

\end{itemize}

As implications, our results can help researchers and practitioners to better understand atoms of confusion in JavaScript. For example, our findings might alert the JavaScript community to avoid writing code with certain atoms of confusion. Moreover, our mixed study might help researchers and practitioners to design program transformation tools to remove atoms of confusion in JavaScript. They might prioritize atoms that cause a lot of confusion and are very common in the repositories, for instance.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% INICIO INTRODUCAO - MARCIO
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Understanding code is an expensive and important task. In real-world software development, it is commonplace for developers to spend more time understanding code than actually writing it~\cite{CitationRequired}.Figuring out what parts of an existing system do by reading its source code is necessary for important software development tasks such as testing, debugging, and adding new functionality to a system.

Developers are often confused while reading unfamiliar code~\cite{Ebert:2021:ESC}. According to Hermans~\cite{ProgrammersBrain}, that confusion can stem from three sources: (i) lack of knowledge, i.e., not knowing what an element in the program does; (ii) lack of information, i.e., not knowing how a program element works; and (iii) lack of processing power, i.e., the inability to combine all the elements in a program and make sense of its execution in one's head. Even small pieces of code can be confusing for developers~\cite{Ajami:2017:SPI,DBLP:conf/sigsoft/GopsteinIYDZYC17} at first glance.

For instance, consider the pair of JavaScript code snippets in Listing~\ref{fig:lst01} (adapted from \textsc{NervJS/taro}). The code snippet on the left-hand side is arguably harder to understand than the one on the right-hand side. The former uses the logical AND operator (\texttt{\&\&}) beyond its lexical meaning~\cite{castor2018} to determine the control flow of the program. Furthermore, the expression has a potential side effect, due to the use of the post-increment operator (\texttt{s++}). The snippet on the right-hand side uses an \texttt{if} statement for control flow and increments the value of variable \texttt{s} using an assignment, making it easier to understand. Notwithstanding, confusion is a direct consequence of the experience of the person who is reading the code. A developer who customarily employs the idiom on the left-hand side may find it easier to understand.

\begin{figure*}[thb]
\noindent\begin{minipage}{.48\textwidth}  
\begin{lstlisting}[language=JavaScript]
res = ps.reduce((s, b) => {
  b && s++
  return s
}
\end{lstlisting}
\end{minipage}\hfill
\begin{minipage}{.48\textwidth}
\begin{lstlisting}[language=JavaScript]
res = ps.reduce((s, b) => {
  if(b) {
    s = s + 1
  }
  return s
}
\end{lstlisting}
\end{minipage}
\captionof{lstlisting}{Code snippet from \textsc{NervJS/taro} project (left-hand side) and an alternative implementation (right-hand size).}\label{fig:lst01}
% On the right-hand side, we present an alternative implementation without atoms of confusion. We could simplify the code on the left-hand side even more, by using \texttt{filter} and \texttt{length} instead of the \texttt{reduce} recursive pattern (e.g., \texttt{res = properties.filter(p => p.length})).}
\end{figure*}

%Line 2 on the left-hand side of the figure contains two atoms of confusion discussed in ~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}: Logic as Control Flow and Post-increment.

Recent work~\cite{DBLP:journals/ese/MedeirosLAAKRG19,DBLP:conf/sigsoft/GopsteinIYDZYC17,Langhout:2021:ACJ,TheEyesDoNotLie} has attempted to elicit and catalog simple code patterns and language constructs that tend to confuse developers when reading them and for which there are less confusing, functionally-equivalent alternatives. These confusing constructs and patterns are called ``atoms of confusion''~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} when it is possible to experimentally ascertain that there is a less confusing alternative (otherwise they are called ``atom candidates''). The snippet on the left-hand side of Figure~\ref{fig:lst01} contains two atoms of confusion discussed in previous work~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} focusing on the \clang language: Logic as Control Flow and Post-increment. Previous work has identified atoms of confusion in two languages, \clang and Java. According to the studies of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms Logic as Control Flow and Post-increment from small \clang code snippets improved the ability of study participants to predict their outcomes by 41\% and 34\%, respectively. For small Java code snippets, Langhout and Aniche~\cite{Langhout:2021:ACJ} report improvements of 53\% and 46.27\%, respectively. These results show that atoms of confusion may be \textbf{challenging} for developers using these languages. In addition, a study~\cite{DBLP:conf/msr/GopsteinZFC18} with 14 large-scale open source projects written in \clang found that 4.38\% of the lines of code have an atom and their presence has a strong correlation with bug fixing commits and long code comments. These results show that atoms of confusion are \textbf{prevalent} in large systems. They are also \textbf{relevant}, even for experienced software developers. 
%According to the experiments of Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17}, removing the atoms Logic as Control Flow and Post-increment improved the accuracy of program understanding tasks in 41\% and 34\%, respectively.

%factors such as the wrong choice of names for variables and methods~\cite{avidan:icpc2017} and the use of uncommon language constructs~\cite{VEM2018} may confused developers might hinder developers from understanding a piece of code. Besides that, programming languages can also have specific idioms and constructs that may lead to misunderstandings. That is the case of atoms of confusion: small pieces of code that prevent developers from program comprehension, even though functionally equivalent alternatives exist and are easier to understand~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17, DBLP:conf/msr/GopsteinZFC18}.

The goal of this paper is twofold. First, it aims to investigate whether atoms of confusion that were identified in Java and C also cause confusion to JavaScript developers. The latter, in spite of the syntactic similarities to Java, is a dynamically-typed language that is arguably more related to Scheme than it is to Java~\cite{Eich:2018:BHJ}. \revised{Previous work has shown that, even for similar languages such as C\# and Java, programmer behavior may differ significantly~\cite{Cabral:2007:EHF,Ebad:2016:ECJ}.}
% This way, there is a knowledge gap with respect to the impact of such atoms in dynamic languages such as JavaScript. In this sense, because 
JavaScript's programming culture differs from that of languages such as \clang and Java, particularly due to its dynamic capabilities and weaker type system. \revised{Furthermore, it includes elements such as array and object destructuring and spread that have no parallel in Java or C.
At the same time, the language is not so different that many} previously identified atoms cannot be represented in it. Second, the paper aims to verify whether other code patterns, not investigated in previous work, also tend to confuse JavaScript developers. Some of these code patterns are specific to JavaScript, e.g., object destructuring. 
% \castor{The previous sentence is really bad.} 
%Second, we want to analyze whether these atoms also cause confusion when we control for developer experience. As mentioned before, experience seems to play a significant role in whether a certain construct or code pattern can be considered an atom of confusion or not. Gopstein et al.~\cite{DBLP:conf/sigsoft/GopsteinIYDZYC17} report that, in their studies, experience is positively correlated with better results. In this paper, we aim to use an experimental design that controls for experience, so as to have greater confidence in our results. 
%, we may find that idioms and constructs that are atoms of confusion in those languages are not particularly difficult to understand for JavaScript programmers and vice-versa. 
%Also, besides being the most widely used language for web development and the most often adopted language on GitHub~\cite{BugsJS:ICST}, 
%JavaScript contains features that combined make the source code fertile ground for errors~\cite{BugsJS:ICST,ClientSideBugsJS:TSE} and confusion~\cite{JSGoodParts:book}.\castor{I think we should only say this if we explicitly point out examples of such features}. 
%, such as weakly typed language~\cite{DynamicBehaviorJS:PLDI}, object-oriented, functional,
% and imperative constructs~\cite{JS-CodeSmells:SCAM}, and features such as run-time
% evaluation~\cite{BugsJS:ICST,ClientSideBugsJS:TSE}, and asynchronous operations.

\revised{We present the results of a mixed-method research effort based on a repository mining study, two experiments, and a set of interviews. 
%, and a mining software repositories study.
The repository mining study measures the prevalence of the atom candidates and helps us motivate the importance of studying these code patterns in more detail. We mined popular JavaScript projects from GitHub looking for the atom candidates that were investigated in both experiments. We found out 
%We found out that atoms frequently appear in practice. Finally, 
%our MSR effort reveals 
that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems.} 
The two experiments ask the participants to predict the output of small code snippets, where some of them have atom candidates and others do not. These two experiments were designed and conducted independently, by different researchers, with different samples, and complementary methodologies; one uses a repeated measures design~\cite{Keselman:2001:ARM} whereas the other uses a Latin square design~\cite{Hunter-Experimenters}.  
Both our experiments suggest that two code patterns that have been previously observed to confuse C programmers also confuse JavaScript programmers (with statistical significance, after correction): the comma operator and assignments being used as values. In addition, pre- and post-increment operators, when used as values, omitted curly braces, changes of literal encoding, and type conversions have led to confusion in at least one of the two experiments. For all these cases effect sizes were either medium or high. In addition, we have discovered that object destructuring and automatic semicolon insertion are potential JavaScript-specific atoms that have not been discussed in previous work.

To learn more about the misunderstandings in the snippets used in the experiments, we also interviewed 15 experienced professional developers. Participants preferred the code snippets without the atoms in 70\% of the cases. The frequency with which the atoms appear \emph{in the wild} might reveal cases where developers should be more or less careful when writing JavaScript code.

%% The number of correct answers for the \emph{clean version} of the code snippets for 7 out of \na atom candidates is at least 15\% higher than the corresponding \emph{confusing version} of the code snippet (i.e., with an atom candidate). For 5 of these atoms, the difference is statistically significant (accounting for p-value correction). Effect size is large for 4 atoms and medium for 1 of them. 
%% The presence of an atom candidate might also affect the time necessary for submitting a correct answer. For instance, the presence of the Comma Operator atom increases the average time to predict the output by 76.23\%, though the outputs of snippets containing the Pre-Increment atom were correctly predicted 38.19\% faster.
%% To learn more about the misunderstandings in the snippets used in the survey, we also interviewed 15 experienced professional developers. Participants preferred the code snippets without the atoms in 70\% of the cases, therefore corroborating the survey findings. The participants also discussed other idioms and constructs that might lead developers to misunderstand JavaScript code.

%% The frequency in which the atoms appear \emph{in the wild} might reveal cases where developers should be more or less careful when writing JavaScript code. To understand the incidence of atoms, we mined popular JavaScript projects from GitHub and found that atoms frequently appear in practice. 
%% Our MSR effort reveals that atom candidates are frequent and used intensively in 72 popular open-source JavaScript systems. Four atom candidates appear in 90\% of the analyzed projects and two of them, Implicit Predicate and Ternary Operator, occur more than once for every 100 lines of code in the dataset. 
%Indeed, four atom candidates are frequently used in JavaScript code---the Implicit Predicate atom, for instance, occurs 19.89 times per 1,000 lines of code.

%\castor{I would rather remove the following. We don't even talk about the "library and infrastructure".}
%Altogether, the main contributions of this paper are:

%\begin{itemize}

%\item A mixed-method research effort---based on two experiments, interviews, and a software
%  repository mining study---about the impact of atoms of confusion on JavaScript programs.
  % (a dynamic language), controlling for subject experience.

%  \item A comparison of the impact and prevalence of atoms of confusion in
%    JavaScript with what has been previously reported for other languages
%    (\clang, \cpplang, and Java).
%    \item A library and infrastructure to mine atoms of confusion using 
%    code queries (implemented in the .QL language). This library and 
%    infrastructure are publicly available and might help researchers and 
%    practitioners to automate program analysis at a large scale.

%\end{itemize}

%As implications, our results can help researchers and practitioners to better understand atoms of confusion in JavaScript. For example, our findings might alert the JavaScript community to avoid writing code with certain atoms of confusion. Moreover, our mixed study might help researchers and practitioners to design program transformation tools to remove atoms of confusion in JavaScript.

%They might prioritize atoms that cause a lot of confusion and are very common in the repositories, for instance.

### 2023-01-11 main.tex
improving threats to validity
Added file response_reviewer_1_2nd_round.md that addresses all the comments made by reviewer 1. I have also changed the order of the studies and made the appropriate adaptations for the MSR study to appear first, before the others and as a motivation for them.

\newcommand{\rqd}{What is the frequency of occurrence of atoms of confusion in practice (i.e., in open-source JavaScript projects)?}

@@

\lstdefinelanguage{CodeQL}[]{Java}{
  morekeywords={select, from, where, not, and}
}
\newcommand{\na}{10\xspace}

@@

\input{introduction}
\input{background}
\input{methodology}
\input{s01}
\input{s02}
\input{s03}
\input{s04}
\input{discussion}
\input{threats}
\input{conclusion}

>>>>

\newcommand{\rqd}{What is the frequency of occurrence of atom candidates in practice (i.e., in open-source JavaScript projects)?}

@@

\lstdefinelanguage{CodeQL}[]{Java}{
  morekeywords={predicate, select, from, where, not, and, or}
}
\newcommand{\na}{10\xspace}

@@

\input{introduction}
\input{background}
\input{methodology}
\input{s04}
\input{s01}
\input{s02}
\input{s03}
\input{discussion}
\input{threats}
\input{conclusion}

### 2023-01-11 methodology.tex
Added file response_reviewer_1_2nd_round.md that addresses all the comments made by reviewer 1. I have also changed the order of the studies and made the appropriate adaptations for the MSR study to appear first, before the others and as a motivation for them.

\begin{enumerate}[(RQ1)]
\item \rqa 
\item \rqb
% \item \rqc  
\item \rqd
\end{enumerate}

By answering these research questions, we can either generalize or refute the findings about atoms of confusion already discussed in the literature
(goal of research questions (RQ1) and (RQ2)). 
In addition, answering the third question allows us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice. Answering these research questions also lays the foundations for the implementation of tools that can automatically transform code into cleaner versions.
%---though we postpone these results to future research work.

@@

These two experiments explore the impact of atom candidates on understanding JavaScript code (goal of research question RQ1). They investigate whether or not programs that contain atom candidates tend to produce more misunderstanding for programmers trying to predict their outcome.% We leverage two independently-developed and conducted experiments with different designs to more  
We also conduct a set of interviews with developers (Section~\ref{sec:s03}) to contrast the quantitative results of the experiments with their preferences and opinions (thus addressing RQ2). Finally, we perform a repository mining study (Section~\ref{sec:s04}) that investigates the prevalence of some of these atom candidates in large scale, professionally-developed JavaScript software.

>>>>

\begin{enumerate}[(RQ1)]
\item \revised{\rqd}
\item \rqa 
\item \rqb
% \item \rqc  
\end{enumerate}

\revised{Answering RQ1 enables us to better motivate the other two questions by showing that studying these code patterns has practical relevance.}
Futhermore, by addressing research questions RQ2 and RQ3, we can either generalize or refute the findings about atoms of confusion already discussed in the literature. 
In addition, answering the third question allows us to enrich existing catalogs about atoms of confusion and discuss how often they occur in practice. Answering these research questions also lays the foundations for the implementation of tools that can automatically transform code into cleaner versions.
%---though we postpone these results to future research work.

@@

\revised{We perform a repository mining study (Section~\ref{sec:s04}) that investigates the prevalence of some of these atom candidates in large scale, professionally-developed JavaScript software. This helps us motivate the remaining three studies.}
These two experiments explore the impact of atom candidates on understanding JavaScript code (goal of research question RQ1). They investigate whether or not programs that contain atom candidates tend to produce more misunderstanding for programmers trying to predict their outcome.% We leverage two independently-developed and conducted experiments with different designs to more  
We also conduct a set of interviews with developers (Section~\ref{sec:s03}) to contrast the quantitative results of the experiments with their preferences and opinions (thus addressing RQ2). %Finally, we perform a repository mining study (Section~\ref{sec:s04}) that investigates the prevalence of some of these atom candidates in large scale, professionally-developed JavaScript software.

### 2023-01-11 s01.tex
Added file response_reviewer_1_2nd_round.md that addresses all the comments made by reviewer 1. I have also changed the order of the studies and made the appropriate adaptations for the MSR study to appear first, before the others and as a motivation for them.

@@

\begin{table}[th!]
\caption{Hypotheses Testing (correctness). Asterisks ($^{*}$) indicate a statistically significant difference. Using Bonferroni correction, a p-value is considered statistically significant if it is lower than 0.0021.}
 \centering
 {\scriptsize 
 \begin{tabular}{lrr}
   \toprule
Atom candidate & McNemar test &      Odds Ratio \\
\midrule
Object Destructuring      &   \textbf{$<$ 0.0001*} & $\infty$ \\
Type Conversion           &   \textbf{$<$ 0.0001*} & 12.0 \\
Change of Literal Encoding   &   \textbf{$<$ 0.0001*} & 28.0 \\
Comma Operator            &   \textbf{$<$ 0.0001*} & 11.5 \\
Lack of Indentation No Braces     &   \textbf{$<$ 0.0001*} &  11.0 \\
Assignment As Value       &   \textbf{$<$ 0.0001*} &  9.3 \\
Pre-Increment              &  \textbf{0.0003*} &    16.0 \\
Omitted Curly Braces      &  \textbf{0.0009*} &  6.3 \\
Array Destructuring       &  0.0023 &     7.5 \\
Post-Increment              &  0.0225 &     5.5 \\
Array Spread              &  0.0490 &    3.3 \\
Object Spread             &  0.0636 &     2.8 \\
Property Access           &  0.0636 &     2.8 \\
Infix Operator Precedence &  0.1796 &     2.5 \\
Arrow Function            &  0.2891 &     3.0 \\
Arithmetic As Logic       &  0.3323 &  1.8 \\
Automated Semicolon Insertion  &  0.3915 &  1.4 \\
Logic As Control Flow     &  0.4421 &  1.5 \\
%Constant Variables        &  0.5000 &     0.0 \\
Implicit Predicate        &  0.6900 &  1.3 \\
Lack of Indentation With Braces   &  0.7266 &  1.7 \\
Conditional Operator      &  0.7905 &  1.3 \\
Dead Unreachable Repeated &  1.0000 &     1.5 \\
Repurposed Variables      &   1.0000 &     1.0 \\
\bottomrule
\end{tabular}
}
\label{tab:repeated:misunderstanding}
\end{table}

>>>>

\revised{The results of the repository mining study show that some of the atom candidates we aim to study occur very frequently. Taking that as a starting point, we conduct two experiments and an interview study to gauge their potential to cause confusion. Furthermore, since we are focusing on JavaScript code, we also analyze a number of atom candidates that are specific to this language in one of the experiments.}

@@

\begin{table}[th!]
\caption{Hypotheses Testing (correctness). Asterisks ($^{*}$) indicate a statistically significant difference. Using Bonferroni correction, a p-value is considered statistically significant if it is lower than 0.0021.}
 \centering
 {\scriptsize 
 \begin{tabular}{lrr}
   \toprule
Atom candidate & McNemar test &      Odds Ratio \\
\midrule
Object Destructuring      &   \textbf{$<$ 0.0001*} & $\infty$ \\
Type Conversion           &   \textbf{$<$ 0.0001*} & 12.00 \\
Change of Literal Encoding   &   \textbf{$<$ 0.0001*} & 28.00 \\
Comma Operator            &   \textbf{$<$ 0.0001*} & 11.50 \\
Lack of Indentation No Braces     &   \textbf{$<$ 0.0001*} &  11.0 \\
Assignment As Value       &   \textbf{$<$ 0.0001*} &  9.30 \\
Pre-Increment              &  \textbf{0.0003*} &    16.00 \\
Omitted Curly Braces      &  \textbf{0.0009*} &  6.30 \\
Array Destructuring       &  0.0023 &     7.50 \\
Post-Increment              &  0.0225 &     5.50 \\
Array Spread              &  0.0490 &    3.30 \\
Object Spread             &  0.0636 &     2.80 \\
Property Access           &  0.0636 &     2.80 \\
Infix Operator Precedence &  0.1796 &     2.50 \\
Arrow Function            &  0.2891 &     3.00 \\
Arithmetic As Logic       &  0.3323 &  1.80 \\
Automated Semicolon Insertion  &  0.3915 &  1.40 \\
Logic As Control Flow     &  0.4421 &  1.50 \\
%Constant Variables        &  0.5000 &     0.0 \\
Implicit Predicate        &  0.6900 &  1.30 \\
Lack of Indentation With Braces   &  0.7266 &  1.70 \\
Conditional Operator      &  0.7905 &  1.30 \\
Dead Unreachable Repeated &  1.0000 &     1.50 \\
Repurposed Variables      &   1.0000 &     1.00 \\
\bottomrule
\end{tabular}
}
\label{tab:repeated:misunderstanding}
\end{table}

### 2023-01-11 s02.tex
Added file response_reviewer_1_2nd_round.md that addresses all the comments made by reviewer 1. I have also changed the order of the studies and made the appropriate adaptations for the MSR study to appear first, before the others and as a motivation for them.

{\bf Exploratory Data Analysis.}
Table \ref{tab:difference-time-taken} shows the average time necessary for the participants to give a correct answer about the expected outcomes of a code snippet, considering both obfuscated and clean versions. We do not consider cases where the participants made mistakes. Seven atom candidates required more time from participants to predict a correct response. For these seven atom candidates, developers take at least 5.90\% less time on average to find a correct answer when considering the clean version of a code snippet. In the extreme case (atom candidate Comma Operator), the participants take 76.23\% less
time on average to find the correct answer for the clean version of the code snippets. 
Not all atom candidates, though, require less time for the participants to predict the answer. In fact, for  Arithmetic as Logic and Pre-Increment, the time to give a correct answer for the clean versions was   
29.06\% and 38.19\% smaller than for the obfuscated versions---we did not confirm these atom candidates as atoms of confusion in the previous section.

>>>>

{\bf Exploratory Data Analysis.}
Table \ref{tab:difference-time-taken} shows the average time necessary for the participants to give a correct answer about the expected outcomes of a code snippet, considering both obfuscated and clean versions. We do not consider cases where the participants made mistakes. Seven atom candidates required more time from participants to predict a correct response. For these eight atom candidates, developers take at least 5.90\% less time on average to find a correct answer when considering the clean version of a code snippet. In the extreme case (atom candidate Comma Operator), the participants take 76.23\% less
time on average to find the correct answer for the clean version of the code snippets. 
Not all atom candidates, though, require less time for the participants to predict the answer. In fact, for  Arithmetic as Logic and Pre-Increment, the time to give a correct answer for the clean versions was   
29.06\% and 38.19\% smaller than for the obfuscated versions---we did not confirm these atom candidates as atoms of confusion in the previous section.
